Seismic parameter estimation and the Canadian crust

### 2013-10-10 paper/draftBP.tex
add Nik and MB edits

We now consider the average crustal structure across Canada. We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method. The resulting profiles contain energy at depths corresponding to major velocity discontinuities. By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components. Figure \ref{fig:conradPCA} contains the first five singular vectors or modes normalized to unit magnitude that account for 78\% of the variance in the profiles. Higher modes correspond to diminishing contributions to total variance. Each vector is normalized to unit magnitude. The first mode (blue) is effectively the average depth profile across all stations and captures 50\% of the variance in the dataset. It features a maximum near 38.7 km that may be taken as an average Moho depth. The second and third modes are dominated by a high amplitude shallow structures (e.g. sedimentary layers) within the top 10 km and Moho variability between 30 and 46 km . The fourth and fifth modes account for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10 km to 20 km range. All 5 modes display markedly lower energy within a depth window roughly between 20 km and 30 km.

@@

Seismic reflectivity of the continental crust is a valuable constraint on the processes governing crustal evolution. Research into the Mohoveric discontinuity and crustal thickness is ongoing and providing new insights into ...

1. introductory paragraph along the lines of the one you did for crustal composition, identifying the
importance of 1-D crustal structure for models of evolution. Introduce Conrad discontinuity and
general seismic reflection subdivision of crust into more transparent upper half and reflective
lower half, ie, summarize basic ideas from Brown EOS article on Conrad and Brocher and Mooney
GJRAS article

2. cite station sorted figure + perhaps normalized stack figure (as per above) and discuss where
energy lies, most pronounced near surface and Moho, lower energy levels in bottom 1/3

3. draw association between bottom 1/3 of low teleseismic reflectivity and high lower crustal  (seismic
reflection) reflectivity, explain effect of frequency in different response

4. discuss evidence for Conrad as distinct discontinuity (doesn't appear to be) versus the more
general subdivision of upper and lower crusts with distinct properties.

Mean crustal composition is an important parameter characterizing continental crust that should not be confused with actual lithological or mineralogical state of the crust. Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time. Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003]. After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal composition is equivalent to the average convergent-margin andesite. Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast. Hence this model cannot account for the intermediate composition of post-Archean crust. Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995]. Unlike these latter studies, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et. al. [1981] that combines $V_P/V_S$ from teleseismic measurements with contraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.

@@

In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a plot $V_P$ versus $V_S$. Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Christensen [1996], all quoted at average crustal pressures of 600 MPa. Dashed lines delimit isocontours in $V_P/V_S$. The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province [Percival and West, 1994]. The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite facies (hence the inclusion of hornblendite and amphibolite) and early plutonic structures comprised largely of tonalites. Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons [Rudnick, 1995].

Seismic parameter estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two candidate end-member lithologies: granite geniss and diorite. This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition [Rudnick and Fountain, 1995; Christensen and Mooney [1995]. The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more tonalitic in average composition. All three provinces cluster about the average value for the Canadian Shield. The Grenville Province, in contrast, displays markedly lower $V_S$ and higher $V_P/V_S$ implying a significant difference in mean composition. The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study [Eaton et. al., 2005] that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition corresponding composition that is more mafic than other Shield provinces. This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally [Eaton et. al., 2005; Rivers, 2008]. However, many stations in the south-east of the province also display $V_P/V_S$ ratios greater than 1.78, similar to the high values common in the adjacent St. Lawrence Platform (where the geometric mean seismic ratio is 1.79). Moreover, and more importantly, the average $V_P$ values for the Grenville province (a region well sampled by refraction profiles) requires low $V_S$ values that are inconsistent with a more mafic composition. Alternative explanations that allow for both high $V_P/V_S$ and reduced $V_P$, $V_S$ include the possibility that Grenville crust contains higher fluid concentrations than that of other tectonic provinces. (NIK?? any others?)

\begin{table}
  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} }
    \hline
    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work \\
    \hline
    $SiO_2$   & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 \\
    $TiO_2$   & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 \\
    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 \\
    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 \\
    $FeO$     & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 \\
    $MnO$     & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
    $MgO$     & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 \\
    $CaO$     & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 \\
    $Na_2O$   & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 \\
    $K_2O$    & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 \\
    $H_2O$    & --- & --- & 0.8 & --- & 1.1 \\
    $P_2O_5$  & --- & 0.2 & 0.2 & 0.1 & 0.2 \\
    $CO_2$    & --- & --- & --- & --- & 0.1 \\
    \hline
  \end{tabular}
  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Fountain [1995] and Rudnick and Gao [2003] use a combined $Fe$ quantity and is listed under $FeO$}
\label{table:composition}

@@

Since our average seismic parameter estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (\ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite. The corresponding chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies. The chemical estimates produced in our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $Fe$ and $Mg$ content than these previous studies. It is noteworthy that that the two studies involving greatest reliance on seismic attributes, ours and the {\it P}-wave velocity study by Smithson [1978], bear the closest similarity. Figure \ref{fig:composition} displays the major crustal element contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition. The gray-shaded region indicates the $\pm 30\%$ range from our reference average Canadian crustal model.  Also included in this figure is the our average crustal composition estimate for the Canadian Shield. Its values closely track the global crustal estimates of Christensen \& Mooney [1995] and Smithson [1978]. These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss. In contrast, the Canadian Shield crust aligns more closely with these previous global estimates and is represented as 68\% dioritic and 32\% granitic using the two end-member model.

@@

@@

@@

@@

@@

@@

Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.
Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.

Kennet, B. L. N., (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.

@@

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{discontinuity}
  \caption{}
  \label{fig:disc}
\end{figure}

We now consider the average crustal structure across Canada. We produce a depth profile for each station by stacking receiver functions along traveltime curves corresponding to a range of values in $H$ (0 to 50 km in increments of 100 metres) using the estimates for $R$ computed via the ZK + semblance method. The resulting profiles contain energy at depths corresponding to major velocity discontinuities and are shown in figure \ref{fig:disc} arranged by increasing crustal thickness. By assembling all depth profiles into a matrix, we may decompose the data via SVD into their principal components. Figure \ref{fig:conradPCA} contains the first five singular vectors or modes normalized to unit magnitude that account for 78\% of the variance in the profiles. Higher modes correspond to diminishing contributions to total variance. Each vector is normalized to unit magnitude. The first mode (blue) is effectively the average depth profile across all stations and captures 50\% of the variance in the dataset. It features a maximum near 38.7 km that may be taken as an average Moho depth. The second and third modes are dominated by a high amplitude shallow structures (e.g. sedimentary layers) within the top 10 km and Moho variability between 30 and 46 km . The fourth and fifth modes account for a combined 11\% of the variance in the dataset, contain energy at upper-mid crustal depths in the 10 km to 20 km range. All 5 modes display markedly lower energy within a depth window roughly between 20 km and 30 km.

@@

\section{Discussion}

Seismic refraction and reflection studies of continental crust have afforded valuable constraints on the processes governing crustal evolution. Early efforts to characterize continental crust invoked a 2-layer, slab model comprising lower velocity upper crust and a higher velocity lower crust [Conrad, 1925; Jeffreys, 1926]. Evidence supporting this model was originally presented by Conrad [1925] who cited observations of P arrivals propagating along a mid-crustal interface similar to but shallower than the deeper crust-mantle interface imaged by Andrija Mohorovičić in 1909 [Litak and Brown, 1989]. Interpretations of seismic data invoking both the crustal slab model and the Conrad Discontinuity, as the mid-crustal interface became known, persisted in the literature but were increasingly criticized as being too simplistic, in the case of the slab model, or, for the Conrad Discontinuity, lacking consistent evidence [Litak and Brown, 1989]. Technological improvements in reflection seismology allowed researchers to further argue for a complex heterogeneous crust and propose models wherein smooth random lateral inhomogeneities are superimposed on simple velocity gradient models to reproduce the range of seismic observations [Litak and Brown, 1989; Mereu and Ojo, 1981].

Despite diminishing support in recent years for the Conrad Discontinuity as a universal crustal feature, the division of the crust into seismically distinguishable upper and lower components remains a common assertion. In their review, Mooney and Brocher [1987] argued for a model of the continental crust separated intoa seismically reflective lower crust overlain by transparent upper crust. Their physical interpretation is that of a poorly laminated upper crust dominated by features of short length scale that contrasts with a lower crust composed of thin ($\sim$ 100 m) lamellae of significantly longer horizontal length scale. The inference of a laminated lower crust has since found support in studies of regions with extended or thicker, warmer crust [e.g. Ito, 1998; Mayer, 1996; Meissner et. al., 2006].

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{discontinuityInterp}
 \caption{Left: Vertical station depth profiles interpolated to a common Moho. Right: Normalized stack of depth profiles, grey line shows mean}
 \label{fig:discInterp}
\end{figure}

The lower energy in the 20-30 km depth window noted in Section \ref{section:discontinuities} for the station profiles given in figure \ref{fig:disc} provides some supporting evidence for seismically distinct upper and lower crustal layers. To better compare results across stations, we normalize each depth profile such that the Moho falls on a common ordinate (Figure \ref{fig:discInterp}, left) and stack the absolute values (Figure \ref{fig:discInterp}, right). The stacked signal has highest amplitude near the surface, where energy is associated with sedimentary layers, and at the Moho. There is a steady and marked decrease in energy over the bottom third of the profile that cannot be ascribed to sidelobe structure associated with the Moho pulse. This observation is ostensibly at odds with the Mooney and Brocher [1987] high reflectivity lower crustal model. Moreover, reflection profiles from {\sc Lithoprobe} transects [Hammer et. al., 2010] frequently display significant reflected energy through the crust with some transects including the Western Superior transect and the Cassiar in the Western Cordillera displaying pronounced pronounced reflectivity within the lower crust. We reconcile the teleseismic and active source observations of transparent versus reflective lower crust by considering the scale of vertical heterogeneity to which the methods are sensitive. Teleseismic body waves include both direct and reflected conversions that can originate from layering with velocity gradients over scale lengths of the order of $\lambda/2$ and $\lambda/4$, respectively, where $\lambda$ is the {\it P}-wavelength [Bostock, 1999]. For 1 Hz teleseismic waves, crustal layering with scale lengths of 1000-2000 m should be resolvable. Higher frequency active source data (10-40 Hz) employing {\it P}-wave sources are sensitive to finer scale vertical structure with scales of 75-300 m. We infer then that the vertical scales of lower crustal reflectivity are limited to scale lengths under 100 m yielding a lower crustal structure that is relatively transparent to teleseismic waves but can yield strongly reflective signatures to near-vertical, active-source seismic profiling.  Some authors, for example Mayer et .al. [1996] in their study of the Rhine Graben, have employed the term Conrad discontinuity to to refer to an interface or transition between transparent upper crust and reflective lower crust. It is possible in this context that the higher energy levels between 15 and 20 km on the 5th principal component in Figure 12, is a manifestation of such
signals, but the lower energy levels ($<5\%$) in this mode, would indicate that Conrad Discontinuity so-defined is at best weakly defined within the Canadian crust. The increased transparency of the lower crust would appear to be better defined and would appear to relate to changes in rheology. Studies on crustal strength [Jackson, 2002] and on brittle and ductile models [Burov and Watts, 2006; Meisner and Rabbel, 2006] as well as the widely noted reduced seismicity of the lower crust indicate changes would support this contention, though see [Deichmann, 1991; Handy and Brun, 2004; Simpson, 1999] for competing evidence.

Mean crustal composition is another important parameter characterizing continental crust though it is important to distinguish it from actual lithological or mineralogical state of the crust. Rather, mean crustal composition is a representative average quantity that serves as a constraint on crustal structure and evolution in space and time. Early attempts at estimating mean crustal composition were largely performed through extrapolations based on the surface sampling of upper crustal rocks [Rudnick and Gao, 2003]. After the plate-tectonic revolution, Taylor [1967, 1977], proposed the 'island arc' or 'andesite' model which posited that bulk crustal chemical composition is equivalent to the average convergent-margin andesite. Several problems with this theory subsequently emerged; for example, average surficial island arc composition is more basaltic than andesitic and any crustal growth via basaltic basal accretion will only serve to magnify this compositional contrast. Hence this model cannot account for the intermediate composition of post-Archean crust. Recognition of these shortcomings led to the analysis of data from crustal refraction seismic surveys to interrogate middle and lower crustal levels and place constraints on their composition [Christensen and Mooney, 1995; Rudnick and Fountain, 1995]. Unlike these latter studies, which also utilize surface and exposed upper-crust geological information, the present work adopts a bulk crustal seismic analysis along the lines of Smithson et. al. [1981] though whereas Smithson utilizes mean crustal $V_P$ this study combines $V_P/V_S$ from teleseismic measurements with constraints on $V_P$ from previous active source studies as presented in the crust 1.0 model.

@@

In Figure \ref{fig:lith} we plot our area-weighted seismic parameter estimates for the entire Canadian data set along with area-weighted average estimates for various tectonic subdivisions on a plot $V_P$ versus $V_S$. Also plotted are values for a selection of mineral assemblages considered to be representative of common crustal rocks from the compilation of laboratory measurements by Christensen [1996], all quoted at average crustal pressures of 600 MPa. Dashed lines delimit isocontours in $V_P/V_S$. The choice of common shield lithologies is motivated by observations along 30km of paleo-depth represented in the Kapuskasing uplift of the Superior province [Percival and West, 1994]. The Kapuskasing uplift comprises Archean granite-greenstone terranes dominated by granitoid and gneissic rocks, and contains supracrustal assemblages generally metamorphosed from greenschist to amphibolite and granulite facies and early plutonic structures comprised largely of tonalites and their metamorphosed equivalents. Tonalites and granodiorites are also part of the TTG (tonalite-trondhjemite-granodiorite) suite which dominates Archaean cratons [Rudnick, 1995].

Seismic parameter estimates for Canada and the Canadian Shield fall, to a first-order approximation, between those of two candidate end-member lithologies: granite geniss and diorite. This intermediate or andesitic composition is in general agreement with other earlier studies of bulk crustal composition [Rudnick and Fountain, 1995; Christensen and Mooney [1995]. The Slave, Superior and Churchill provinces possess comparable seismic parameters implying similar mean crustal compositions, with the Slave province represented as slightly more felsic in average composition. All three provinces cluster about the average value for the Canadian Shield. The Grenville Province, in contrast, displays markedly lower $V_S$ and higher $V_P/V_S$ implying a significant difference in mean composition. The high $V_P/V_S$ estimate, 1.77, is similar to the averaged results from a previous seismic study [Eaton et. al., 2005] that included stations of the central Grenville Orogen, and might be interpreted to indicate that the Grenville province possesses a mean crustal composition corresponding composition that is more mafic than other Shield provinces. This interpretation is consistent with a tectonic model for the the central Grenville Province which suggests that the mid-crust has been completely excised locally [Eaton et. al., 2005; Rivers, 2008]. However, many stations in the south-east of the province also display $V_P/V_S$ ratios greater than 1.78, similar to the high values common in the adjacent St. Lawrence Platform (where the geometric mean seismic ratio is 1.79). Moreover, and more importantly, the average $V_P$ values for the Grenville province (a region well sampled by refraction profiles) requires low $V_S$ values that are inconsistent with a more mafic composition. Alternative explanations that allow for both high $V_P/V_S$ and reduced $V_P$, $V_S$ include the possibility that Grenville crust contains higher fluid concentrations than that of other tectonic provinces. [[Double checkout less quartzite argument, this implies higher Vp ... but the data does not show it ]]

\begin{table}
  \begin{tabular}{ p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm} p{2cm}}
    \hline
    Oxide, wt\% & Smithson [1978] & Rudnick and Fountain [1995] & Christensen and Mooney [1995] & Rudnick and Gao [2003] & This Work: Canada & This Work: Shield\\
    \hline
    $SiO_2$ & 63.0 & 52.6 & 61.7 & 60.6 & 64.3 & 61.9 \\
    $TiO_2$ & 0.7 & 0.8 & 0.9 & 0.7 & 0.6 & 0.7 \\
    $Al_2O_3$ & 15.8 & 16.6 & 14.7 & 15.9 & 15.5 & 15.9 \\
    $Fe_2O_3$ & 2.0 & --- & 1.9 & --- & 1.9 & 2.1 \\
    $FeO$ & 3.4 & 6.6 & 5.1 & 6.7 & 3.3 & 3.9 \\
    $MnO$ & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 & 0.1 \\
    $MgO$ & 2.8 & 4.4 & 3.1 & 4.7 & 2.2 & 2.8 \\
    $CaO$ & 4.6 & 6.4 & 5.7 & 6.4 & 4.3 & 5.1 \\
    $Na_2O$ & 4.0 & 3.2 & 3.6 & 3.1 & 3.6 & 3.6 \\
    $K_2O$ & 2.7 & 1.9 & 2.1 & 1.8 & 2.9 & 2.5 \\
    $H_2O$ & --- & --- & 0.8 & --- & 1.1 & 1.2 \\
    $P_2O_5$ & --- & 0.2 & 0.2 & 0.1 & 0.2 & 0.2 \\
    $CO_2$  & --- & --- & --- & --- & 0.1 & 0.1 \\
    \hline
  \end{tabular}
  \caption{Bulk crustal composition comparison from selected sources. Rudnick and Fountain [1995] and Rudnick and Gao [2003] use a combined $Fe$ quantity and is listed under $FeO$}
\label{table:composition}

@@

Since our average seismic parameter estimates for the entire Canadian crust lie on a line connecting diorite with granite gneiss (\ref{fig:lith}), we may estimate bulk crustal composition in proportion to these lithologies as 51\% diorite and 49\% granite. The lithological estimate for the Canadian Shield is 68\% diorite and 32\% granite. This lithological ratio is applied to the chemical compositions of diorite and granite using chemical averages calculated from 141 diorite and 197 granite samples compiled by Le Maitre [1976] to produce an estimate of bulk crustal composition. The resulting chemical breakdown is provided in table \ref{table:composition} along with previous estimates from global crustal studies. Our analysis reveal a crust with marginally higher average $SiO_2$ composition and lower $Fe$ and $Mg$ content than these previous studies. It is noteworthy that that the two studies involving greatest reliance on seismic attributes, ours and the {\it P}-wave velocity study by Smithson [1978], bear the closest similarity. Figure \ref{fig:composition} displays the major crustal element contributions from the data in table \ref{table:composition} normalized to our average Canadian crustal composition. The gray-shaded region indicates the $\pm 30\%$ range from our reference average Canadian crustal model.  Also included in this figure is the our average crustal composition estimate for the Canadian Shield. Its values closely track the global crustal estimates of Christensen \& Mooney [1995] and Smithson [1978]. These data imply that the extensive platform and orogenic regions of Canada skew the average composition away from the global estimates of previous studies through incorporation of a higher proportion of granite gneiss. In contrast, the Canadian Shield crust aligns more closely with these previous global estimates.

@@

Bostock, M. G. (1999), Seismic waves converted from velocity gradient anomalies in the Earth’s upper mantle, Geophysical Journal International, 138, 747-756.

@@

Burov, E. B., A. B. Watts (2006), The long-term strength of continental lithosphere:" jelly sandwich" or" crème brûlée"?, GSA today, 16, 4.

@@

Conrad, V. (1925), Laufzeitkurven des Tauernbebens vom 28. November 1923, Mitt. Erdb. Komm. Wiener Akad. Wiss., 59, 1–23.

@@

Deichmann, N. (1992), Structural and rheological implications of lower-crustal earthquakes below northern Switzerland, Physics of the earth and planetary interiors, 69, 270-280.

@@

Golub, G. H., M. Heath, G. Wahba (1979), Generalized cross-validation as a method for choosing a good ridge parameter, Technometrics, 21, 215-223.

Hammer, P. T., R. M. Clowes, F. A. Cook, A. J. van der Velden, K. Vasudevan (2010), The Lithoprobe trans-continental lithospheric cross sections: imaging the internal structure of the North American continent, Canadian Journal of Earth Sciences, 47, 821-857.

Handy, M. R.,  J. P. Brun (2004), Seismicity, structure and strength of the continental lithosphere, Earth and Planetary Science Letters, 223, 427-441.

Ito, K. (1999), Seismogenic layer, reflective lower crust, surface heat flow and large inland earthquakes, Tectonophysics, 306, 423-433.

Jackson, J. (2002), Strength of the continental lithosphere: Time to abandon the jelly sandwich?, GSA today, 12, 4-9.

Jeffreys, H. (1926), On compressional waves in two superposed layers, Proc. Cambridge Philos. Soc., 23, 472–481.

Kennet, B. L. N. (1991), The removal of free surface interactions from three-component seismograms, J. Geophys. Int., 104, 153-163.

@@

Litak, R. K., L. D. Brown (1989), A modern perspective on the Conrad discontinuity, Eos, 70, 713-725.

Mayer, G., P. M. Mai, T. Plenefisch, H. Echtler, E. Lüschen, V. Wehrle, B. Müller, K-P. Bonjer, C. Prodehl, K. Fuchs (1997), The deep crust of the Southern Rhine Graben: reflectivity and seismicity as images of dynamic processes, Tectonophysics, 275, 15-40.

Meissner, R., W. Rabbel, H. Kern (2006), Seismic lamination and anisotropy of the lower continental crust, Tectonophysics, 416, 81-99.

Mereu, R. F., S. B. Ojo (1981), The scattering of seismic waves through a crust and upper mantle with random lateral and vertical inhomogeneities, Physics of the Earth and Planetary Interiors, 26, 233-240.

Mooney, W. D. (2012), Personal communication. Compiled GSC active source data for the Canada.

Mooney, W. D., T. M. Brocher (1987) Coincident seismic reflection/refraction studies of the continental lithosphere: A global review, Rev. Geophys., 25, 723 – 742.

@@

Simpson, F. (1999), Stress and seismicity in the lower continental crust: A challenge to simple ductility and implications for electrical conductivity mechanisms. Surveys in Geophysics, 20, 201-227.
